---
title: "Causal ML Review Data: Simple User Guide"
author: "Sharath Chandra Shankar"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 2
    theme: readable
---

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)
library(tidyr)
library(readr)
library(tibble)
```

# 1. Start here: load the review tables

Users do **not** need to manually create the `.RDS` file or worry about how the `.R` build file works.

Just run this code. It pulls the latest `data_loading_module.R` file from GitHub, creates the review tables, and automatically saves `causal_ml_review.rds`.

```{r load-from-github}
source(
  "https://raw.githubusercontent.com/sharashankr/Causal_ML_Literature_Review_Data_Analyses/main/data_loading_module.R",
  encoding = "UTF-8"
)
```

After running the code above, these tables are available in your R session:

| Table | What it contains |
|---|---|
| `papers` | One row per paper with paper details |
| `scores` | One row per paper with framework scores |
| `evidence` | Long table with one row per paper × scoring dimension |
| `review` | A list containing `papers`, `scores`, and `evidence` |

The script also saves a local file called:

```{r rds-note, eval=FALSE}
causal_ml_review.rds
```

# 2. Look at the tables

```{r inspect-tables, eval=FALSE}
# Basic dimensions
dim(papers)
dim(scores)
dim(evidence)

# Column names
names(papers)
names(scores)
names(evidence)

# First few rows
head(papers)
head(scores)
head(evidence)
```

In RStudio, open the tables like this:

```{r view-tables, eval=FALSE}
View(papers)
View(scores)
View(evidence)
```

# 3. List all papers

```{r list-papers, eval=FALSE}
papers %>%
  select(paper_id, first_author, year, cancer_type, title)
```

# 4. Join paper details with scores

Both `papers` and `scores` contain a column called `study_design`.

To avoid confusing `.x` and `.y` columns after the join, rename them before joining.

```{r join-papers-scores, eval=FALSE}
paper_scores <- papers %>%
  rename(study_design_metadata = study_design) %>%
  left_join(
    scores %>% rename(study_design_score = study_design),
    by = "paper_id"
  )

View(paper_scores)
```

Example: select only a few useful columns.

```{r selected-paper-scores, eval=FALSE}
paper_scores %>%
  select(
    paper_id,
    first_author,
    year,
    cancer_type,
    study_design_metadata,
    study_design_score,
    method_type,
    causal_estimand,
    survival_estimand,
    uncertainty_quantification
  )
```

# 5. Make scores long format

The `scores` table is wide. Each scoring dimension is a column.

For joins with `evidence`, it is easier to convert it to long format.

```{r scores-long, eval=FALSE}
scores_long <- scores %>%
  pivot_longer(
    cols = -paper_id,
    names_to = "dimension",
    values_to = "score"
  )

head(scores_long)
```

# 6. Join scores with evidence

This lets you see the numeric score and the evidence excerpt together.

```{r join-scores-evidence, eval=FALSE}
score_evidence <- scores_long %>%
  left_join(
    evidence,
    by = c("paper_id", "dimension")
  )

View(score_evidence)
```

Useful selected version:

```{r selected-score-evidence, eval=FALSE}
score_evidence %>%
  select(
    paper_id,
    dimension,
    score,
    score_assigned,
    excerpt,
    rater_note
  )
```

# 7. Check one paper

Example: see all scores and evidence for one paper.

```{r one-paper, eval=FALSE}
score_evidence %>%
  filter(paper_id == "Ma_2018") %>%
  select(dimension, score, excerpt, rater_note)
```

Change `"Ma_2018"` to any other `paper_id`.

# 8. Check one scoring dimension across papers

Example: compare survival estimand scoring across all papers.

```{r one-dimension, eval=FALSE}
score_evidence %>%
  filter(dimension == "survival_estimand") %>%
  select(paper_id, score, excerpt, rater_note)
```

Other useful dimensions:

```{r useful-dimensions, eval=FALSE}
score_evidence %>%
  filter(dimension == "uncertainty_quantification")

score_evidence %>%
  filter(dimension == "balance_diagnostics")

score_evidence %>%
  filter(dimension == "causal_estimand")
```

# 9. Simple summary tables

Average scores by dimension:

```{r avg-by-dimension, eval=FALSE}
scores_long %>%
  group_by(dimension) %>%
  summarise(
    mean_score = mean(score, na.rm = TRUE),
    n_scored = sum(!is.na(score)),
    .groups = "drop"
  ) %>%
  arrange(mean_score)
```

Number of papers by cancer type:

```{r cancer-counts, eval=FALSE}
papers %>%
  count(cancer_type, sort = TRUE)
```

Number of papers by study design score:

```{r study-design-counts, eval=FALSE}
scores %>%
  count(study_design)
```

# 10. Export tables

Save as CSV:

```{r export-csv, eval=FALSE}
write_csv(papers, "papers.csv")
write_csv(scores, "scores.csv")
write_csv(evidence, "evidence.csv")
write_csv(score_evidence, "score_evidence_joined.csv")
```

Save as TSV:

```{r export-tsv, eval=FALSE}
write_tsv(papers, "papers.tsv")
write_tsv(scores, "scores.tsv")
write_tsv(evidence, "evidence.tsv")
```

# 11. Optional: read the saved RDS in a future session

The GitHub `source()` step automatically saves `causal_ml_review.rds`.

In a future R session, if you already have that RDS file locally, you can read it directly without sourcing GitHub again.

```{r read-rds-future, eval=FALSE}
review <- readRDS("causal_ml_review.rds")

papers   <- review$papers
scores   <- review$scores
evidence <- review$evidence
```

# 12. Common errors

## Error: object `papers` not found

You probably have not run the GitHub source code yet.

Run this first:

```{r error-papers-not-found, eval=FALSE}
source(
  "https://raw.githubusercontent.com/sharashankr/Causal_ML_Literature_Review_Data_Analyses/main/build_20_papers.R",
  encoding = "UTF-8"
)
```

## Error: cannot open the GitHub connection

Check that:

1. You are connected to the internet.
2. The GitHub repository is public.
3. The file still exists at `main/build_20_papers.R`.
4. The URL has `raw.githubusercontent.com`, not `github.com/.../blob/...`.

## Error: column `study_design` does not exist after joining

This happens when both tables had a `study_design` column and R renamed them as `.x` and `.y`.

Use this safer join:

```{r fix-study-design-join, eval=FALSE}
paper_scores <- papers %>%
  rename(study_design_metadata = study_design) %>%
  left_join(
    scores %>% rename(study_design_score = study_design),
    by = "paper_id"
  )
```

# 13. Minimal workflow

Most users only need this:

```{r minimal-workflow, eval=FALSE}
library(dplyr)
library(tidyr)
library(readr)

source(
  "https://raw.githubusercontent.com/sharashankr/Causal_ML_Literature_Review_Data_Analyses/main/data_loading_module.R",
  encoding = "UTF-8"
)

paper_scores <- papers %>%
  rename(study_design_metadata = study_design) %>%
  left_join(
    scores %>% rename(study_design_score = study_design),
    by = "paper_id"
  )

scores_long <- scores %>%
  pivot_longer(
    cols = -paper_id,
    names_to = "dimension",
    values_to = "score"
  )

score_evidence <- scores_long %>%
  left_join(evidence, by = c("paper_id", "dimension"))

View(papers)
View(scores)
View(evidence)
View(paper_scores)
View(score_evidence)
```
